---
title: "p8105_hw1_my2838"
author: "Mia Yu"
date: "2023-09-18"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(moderndive)
```

# Problem 1

### Read and describe the dataset

```{r}
data("early_january_weather")
nrow(early_january_weather) # calculate the rows
ncol(early_january_weather) # calculate the columns 
mean (pull(early_january_weather,temp)) # calculate the mean temperature
```

1. The variables of the dataset are:

`origin`: Name of weather station.

`year`,`day`, `hour`: Recording time.

`temp`: Temperature.

`dewp`: Dewpoint.

`humid`: Humidity.

`wind_dir`: wind direction (in degrees).

`wind_speed`: wind speed (in mph).

`wind_gust`: gust speed (in mph).

`precip`: Precipitation (in inches).

`pressure` :sea level pressure.

`visib`: Visibility (in miles).

`time_hour`: Record date and hour.

Generally, the weather data are measured from EWR, from of January 1 to January 15, 2013.


2. There are 358 rows and 15 columns in the dataset.


3. The mean temperature is 39.58212.

### Make a scatterplot and describe patterns

```{r}
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) + geom_point() # creat the scatterplot
ggsave("P8105_hw1_my2838_1.pdf") # save the plot
```

The plot shows that the temperature is generally rising from Jan 1 to Jan 15, and also the percentage of humidity increased during that time period.


# Problem 2

### Creat the data frame
```{r}
problem2_df = tibble(
  vec_sample = rnorm(10),
  vec_logical = vec_sample > 0,
  vec_character = c("A","B","C","D","E","F","G","H","I","J"),
  vec_factor = factor(rep(c("a","b","c"), times = c(1,3,6))
)
)
```

### Calculate the mean value of each orginal variable

```{r}
mean(pull(problem2_df, vec_sample))
mean(pull(problem2_df, vec_logical))
mean(pull(problem2_df, vec_character))
mean(pull(problem2_df, vec_factor))
```

It *works* when calculating the mean of vec_sample and vec_logical.
But it *doesn't work* when calculating the mean of vec_character and vec_factor.

### Try `as.numeric` function

```{r, eval=FALSE}
mean(as.numeric(pull(problem2_df,vec_logical)))
mean(as.numeric(pull(problem2_df,vec_character)))
mean(as.numeric(pull(problem2_df,vec_factor)))
```

R reads TRUE as 1 and FALSE as 0. When `as.numeric()` function is applied to logical variable, the logical value will contain the numerical values of 0 and 1.

It is not possible to find the mean of a character variable. When `as.numeric()` function is applied to characters, R cannot identified the character data type into numeric data type and so the result is NA.

It is not possible to find the mean of a factor variable. When `as.numeric()` function is applied to factors, R identified each level in numerical values of 1, 2, 3.

